# Code for Yugoslavia Paper

Author: marten walk

## Imports 

```{r}
library(tidyverse)
library(maddison)
library(ggthemes)
```

## Data Cleaning

### Debt

```{r}
debt <- read_csv2("data/debt.csv")
```

Clean

```{r}
colnames(debt) <- c("year", "gross_debt", "foreign_lending", "net_debt", "IMF", "IBRD", "interest_payments", "amortization", "debt_servicing", "debt_servicing_percent")

debt$debt_servicing_percent <- debt$debt_servicing_percent / 10

```

### Trade

```{r}
trade <- read_csv2("data/trade.csv")
```

combine text of type column and category column

```{r}
trade$Type <- paste(trade$Type, trade$Category, sep = "_")
trade$Category <- NULL
trade$Type <- gsub(" ", "_", trade$Type)
```

transpose the data, until now it is Type,1965,1966,..., make it Year, Type_1,Type_2, Type_3,...

```{r}
trade <- trade %>%
  t(.) %>% # transpose so that the years are in the first column
  as.data.frame(.) %>% # convert to data frame
  setNames(.[1,]) %>% # set the first row as the column names
  .[-1,] %>% # remove the first row
  rownames_to_column("year") %>% # year = index col
  mutate_all(as.numeric) %>%
  select(-c(3,5,9,18)) # remove subcategories
```


### Facts

```{r}
facts <- read_csv("data/facts.csv")
```

```{r}
facts <- facts %>% 
  select(-Unit) %>%
  t(.) %>% # transpose so that the years are in the first column
  as.data.frame(.) %>% # convert to data frame
  setNames(.[1,]) %>% # set the first row as the column names
  .[-1,] %>% # remove the first row
  rownames_to_column("year") # year = index col
```

Rename columns

```{r}
facts <- facts %>%
  rename(
    "GNP_factor_dinar" = "Gross national product at factor cost",
    "GNP_pp_dollar" = "Gross national product per head",
    "GFAF_dinar" = "Gross fixed asset formation",
    "GFAF_percent" = "Gross fixed asset formation as percentage of GDP",
    "GNP_agriculture_percent" = "Origin of GDP - Agriculture",
    "GNP_mining_percent" = "Origin of GDP - Mining and manufacturing",
    "GNP_building_percent" = "Origin of GDP - Building",
    "GNP_other_percent" = "Origin of GDP - Other",
    "G_revenue_percent" = "General government revenue as percentage of GDP"
  ) %>%
  select(-c("Collective consumption as percentage of GDP", "Federal revenue as percentage of GDP")) %>%
  #remove all spaces in data
  mutate_all(str_replace_all, " ", "") %>%
  mutate_all(as.numeric)
```

### Exchange Rate

```{r}
exchange <- read_csv("data/exchange_rate.csv")
```

```{r}
exchange <- exchange %>%
  t(.) %>%
  as.data.frame(.) %>%
  rownames_to_column("year") %>%
  setNames(.[1,]) %>%
  mutate_all(as.numeric) %>%
  remove_missing()
```

### Balance of Payments

```{r}
bop <- read_csv("data/bop.csv")
```

clean

```{r}
bop <- bop %>%
  rename_all(str_replace_all, " ", "_") %>%
  rename_all(tolower) %>%
  #rename column 2 
  rename("exports" = "exports,_f.o.b." ) %>%
  rename("imports" = "imports,_c.i.f." ) %>%
  rename("remit" = "emigrant_and_worker_remittances") %>%
  mutate_all(as.numeric) %>%
  mutate(imports = abs(imports)) %>%
  # multiply all by 1000000
  mutate_all(~ . * 1000000) %>%
  mutate(year = year / 1000000)
  #remove duplicate year 1979
```


### Migration & Budget

```{r}
migration <- read_csv("data/migration.csv")
```
```{r}
investment <- read_csv("data/investment.csv")
```
### Wage 
```{r}
wage <- read_csv("data/wage.csv")
```

Calculate real wage
```{r}
wage$real <- wage$wage / wage$col #cost of living = col
wage$real_growth <- wage$real - lag(wage$real)
```

calc 3year moving average

```{r}
wage$moving_avg <- zoo::rollmean(wage$real_growth, 5, fill = NA, align = "right")
```

## Explorative Plotting

Current Account Balance

```{r}
bop %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = trade_balance), color = "black") +
  geom_line(aes(y = current_balance), color = "green") 
  #geom_col(aes(y = remit), fill = "blue")
```

### Exchange Rate

```{r}
exchange %>%
  filter(Year <=1990) %>%
  ggplot(aes(x = Year, y = log(Rate))) +
  geom_line()
```

Plot the yearly change in the exchange rate

```{r}
exchange %>%
  filter(Year <=1987) %>%
  mutate(change = Rate - lag(Rate)) %>%
  ggplot(aes(x = Year, y = change)) +
  geom_line()
```

### Facts 

Plot the GNP per capita

```{r}
ggplot(aes(x = year, y = GNP_pp_dollar), data=facts) +
  geom_line() +
  labs(title = "GNP per capita",
       x = "Year",
       y = "GNP per capita") 
```

Plot the gov revenue as percentage of GDP

```{r}
ggplot(aes(x = year, y = G_revenue_percent), data=facts) +
  geom_line() +
  labs(title = "General government revenue as percentage of GDP",
       x = "Year",
       y = "Percentage of GDP") 
```


### Trade

Plot Total Imports and Exports and the deficit between them

```{r}
trade %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y=IMPORTS_Total, color="red")) +
  geom_line(aes(y=EXPORTS_Total, color="blue")) +
  geom_line(aes(y=IMPORTS_Total - EXPORTS_Total, color="green")) 
```

Plot the composition of imports (all rows that begin with IMPORTS_, except the IMPORTS_Total) over all years

```{r}
trade %>%
  select(1:9) %>%
  pivot_longer(cols = -year, names_to = "type", values_to = "value") %>%
  ggplot(aes(x = year, y = value, fill = type)) +
  geom_area(position = "stack") 
```

make a percentage area plot of these

```{r}
trade %>%
  select(1:9) %>%
  pivot_longer(cols = -year, names_to = "type", values_to = "value") %>%
  group_by(year) %>%
  mutate(percent = value / sum(value) * 100) %>%
  ggplot(aes(x = year, y = percent, fill = type)) +
  geom_area(position = "stack") +
  theme_linedraw() + 
  #put legend on the graph and the corresponding sections
  theme(legend.position = "bottom") 
```

How much of the rise in imports is due to fuels and chemicals?

```{r}
trade %>%
  ggplot(aes(x = year)) +
  geom_area(aes(y = IMPORTS_Mineral_fuels+ + IMPORTS_Chemicals), fill="red") 
```


## Combine the Data

```{r}
comp_data <- maddison::maddison
comp_data <- comp_data %>%
  #filter(countrycode %in% c("GRC", "PRT", "ESP", "YUG")) %>%
  filter(countrycode %in% c("BRA", "MEX", "YUG", "BGR", "POL")) %>%
  filter(year >= 1950 & year <= 1990)
```


```{r}
comb <- comp_data %>%
  filter(countrycode == "YUG") %>%
  select(rgdpnapc, pop, year) %>%
  mutate(gdp_dollar = rgdpnapc * pop*100) %>%
  left_join(trade, by = "year") %>%
  left_join(debt, by = "year") %>%
  left_join(facts %>% select(year, GNP_pp_dollar, GNP_factor_dinar), by = "year") %>%
  left_join(bop %>% select(year, trade_balance, current_balance, remit), by = "year") %>%
  left_join(migration, by="year") %>%
  left_join(investment, by="year") %>%
  left_join(wage %>% select(year, real_growth, moving_avg), by="year")
```

```{r}
comb <- comb %>%
  mutate_at(vars(starts_with("IMPORTS_"), starts_with("EXPORTS_")), ~.*1000000) %>%
  mutate(gnp_dollar_yug = GNP_pp_dollar * pop*1000)
```

do some deflating exercises etc
```{r}
library(priceR)
```

```{r}
comb$
```

## Plotting of Finished Figures

### Figure: Comparison of GDP Countries

Plot the GDP per capita from 1945-1990

```{r}
shapes <- c(0,1,2,5,6)
comp_data %>%
  ggplot(aes(x = year, y = rgdpnapc, color = country, shape = country)) +
  geom_line(size=0.4) +
  geom_point() +
  scale_color_brewer(palette = "Set1")+  
  labs(title = "GDP per capita from 1945-1990",
       x = "Year",
       y = "GDP per capita") +
  theme_tufte() +
  scale_shape_manual(values = shapes) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    )
ggsave("images/GDP_per_capita.png", width = 7, height = 4, dpi = 500)
```

Plot the growth rate of gdp

```{r}
# # growth rate = (gdp(t) - gdp(t-1)) / gdp(t-1)
# comp_data %>%
#   group_by(country) %>%
#   mutate(growth_rate = (rgdpnapc - lag(rgdpnapc)) / lag(rgdpnapc)) %>%
#   mutate(moving_avg = rollmean(growth_rate, 3, fill = NA, align = "right")) %>%
#   ggplot(aes(x = year, y = moving_avg, color = country, shape = country)) +
#   geom_line(size=0.4)
#   geom_point()
```
### Figure: Energy Imports

```{r}
imports_gdp_plot <- comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=IMPORTS_Mineral_fuels/gdp_dollar*100), color="red")  +
    geom_point(aes(y=IMPORTS_Mineral_fuels/gdp_dollar*100), shape=5, color="red")
imports_gdp_plot
```

```{r}
imports_total_plot <- comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=IMPORTS_Mineral_fuels/IMPORTS_Total*100), color="blue") +
    geom_point(aes(y=IMPORTS_Mineral_fuels/IMPORTS_Total*100), shape=5, color="blue") +
    ylim(0,30)
imports_total_plot
```

```{r}
# imports_gdp_plot +
imports_total_plot +
  scale_color_brewer(palette = "Set1")+  
  labs(
    x = "Year",
    # y = "% of GDP",
    y = "% of Imports"
    ) +
  theme_tufte() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    ) +
  xlim(c(1957,1990))

# ggsave("images/imports_gdp.png", width = 4, height = 3, dpi = 500)
ggsave("images/imports_total.png", width = 4, height = 3, dpi = 500)
```



### Figure: Migration

Plot the number of migrants and net migration

```{r}
ggplot(migration, aes(x=year))+
  geom_line(aes(y=nr_emigration/1000,color="Nr. of Emigrants")) +
  geom_point(aes(y=nr_emigration/1000), color="blue", shape=5) +
  geom_line(aes(y=net/1000, color="Net Migration"))+
  geom_point(aes(y=net/1000), color="red", shape=6) +
  #styling
  scale_color_brewer(palette = "Set1")+  
  labs(
    title = "Yugoslav Migration 1968-1985",
    x = "Year",
    y = "Migrants in Thousands",
    ) +
  geom_hline(yintercept=0, color="grey")+
  theme_tufte() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    ) +
  scale_y_continuous(n.breaks=6)+
  xlim(c(1967,1985)) 
ggsave("images/migration.png", width=7, height=4, dpi=500)
```


### Figure: Remittances

```{r}
remit_total_plot <- comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=remit/1000000), color="orange") +  #absolute
    geom_point(aes(y=remit/1000000), color="orange", shape=5)
remit_total_plot
```

```{r}
remit_rel_plot <- comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=remit/gdp_dollar*100), color="blue") +#relative to GDP
    geom_point(aes(y=remit/gdp_dollar*100), color="blue", shape=5) +
    ylim(0,8)
remit_rel_plot
```

sinking share of worker remittances, absolutely numbers down as well as relative

```{r}
#remit_total_plot +
remit_rel_plot + 
  scale_color_brewer(palette = "Set1")+  
  labs(
    x = "Year",
    # y = "Remittances (Million $, current prices) ",
    y = "% of GDP"
    ) +
  theme_tufte() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    ) +
  xlim(c(1960,1990))

# ggsave("images/remit_total.png", width = 4, height = 3, dpi = 500)
ggsave("images/remit_rel.png", width = 4, height = 3, dpi = 500)
```


### Figure: Debt (Gross and relative)

```{r}
comb %>%
  filter(year > 1960) %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=debt_servicing*1000000/gdp_dollar*100, color="Debt Servicing")) + 
    geom_point(aes(y=debt_servicing*1000000/gdp_dollar*100, color="Debt Servicing"), shape=6) +
    geom_line(aes(y=gross_debt*1000000/gdp_dollar *100, color="Gross Debt")) +
    geom_point(aes(y=gross_debt*1000000/gdp_dollar *100, color="Gross Debt"), shape=5) +
    #styling
    scale_color_brewer(palette = "Set1") +
    labs(
      x = "Year",
      y = "% of GDP"
      ) +
    theme_tufte() +
    theme(
      legend.position = "bottom",
      legend.title = element_blank(),
      panel.border = element_rect(fill = NA, color = "black"),
      ) +
    xlim(c(1960,1986)) +
    scale_y_continuous(n.breaks=7) 
ggsave("images/debt.png", width=7, height=3, dpi=500)
```


### Figure: Trade Balance

Trade Balance

```{r}
bop %>%
  ggplot(aes(x = year)) +
  geom_col(aes(y = -exports/1e6, color="Exports"), fill="red", size=0, alpha=0.6) + 
  geom_col(aes(y = imports/1e6, color="Imports"), fill="blue", size=0, alpha=0.6) +
  geom_line(aes(y = trade_balance/1e6), color="black") +
  geom_point(aes(y = trade_balance/1e6), color="black", shape=5) +
  #styling
  labs(title = "Trade Balance, 1960-1990",
       x = "Year",
       y = "Flows (Million $)"
       ) +
  theme_tufte() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black")
    ) +
  scale_x_continuous(n.breaks=7) +
  scale_y_continuous(n.breaks=6) +
  geom_hline(yintercept=0, color="grey")
ggsave("images/trade_balance.png", width = 7, height = 4, dpi = 500)
```

### Figure: Inv and GDP

```{r}
ggplot(aes(x=year, y=(inv_otal / social_product)*100), data=investment) +
  geom_line() +
  geom_point(shape=5) +
  theme_tufte() +
  labs(title = "Gross fixed Investment and Social Product",
       x = "Year",
       y = "% of SP"
       ) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black")
    ) +
  xlim(1969,1987)
ggsave("images/inv_sp.png", height=3, width=7, dpi=500)
```

### Figure: Real Wage Growth

Plot the real wage growth and the 5-year moving average of the real wage growth

```{r}
ggplot(aes(x = year), data=comb) +
  geom_point(aes(y= real_growth, color="yearly"), shape=3) +
  geom_line(aes(y= moving_avg, color="5-year mov. avg")) +
  geom_point(aes(y= moving_avg), shape=5, color = "red") +
  labs(title = "",
       x = "Year",
       y = "Growth") +
  geom_hline(yintercept=0, color="grey") +
  theme_tufte() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    ) +
  scale_y_continuous(labels = scales::percent_format(scale = 100)) +
  xlim(1970,1990)
ggsave("images/real_wage.png", width = 7, height = 4, dpi = 500)
```

## Additional Tests

do some deflating exercises etc

```{r}
library(priceR)
```

```{r}
bop$adjusted_remit <- adjust_for_inflation(bop$year, bop$remit, "US", to_date = 2011)
```

```{r}
df_90 <- readstata13::read.dta13("https://www.rug.nl/ggdc/historicaldevelopment/maddison/data/mpd2018_1990bm.dta")
```

```{r}
df_90 %>%
  filter(countrycode == "YUG") %>%
  select(gdppc, year) %>%
  filter(year >= 1960)
```
```{r}
comp_data %>%
  filter(countrycode == "YUG") %>%
  select(cgdppc, year) %>%
  filter(year >= 1960)
```


