---
format: html
---

Economic Downfall of Yugoslavia - economic inevitability or political failure?

## Imports

```{r}
library(tidyverse)
library(maddison)
library(ggthemes)
#library(zoo)
```

## Comparison with other Countries

```{r}
comp_data <- maddison::maddison
```

Filter for Greece, Portugal, Spain and Yugoslavia

```{r}
comp_data <- comp_data %>%
  #filter(countrycode %in% c("GRC", "PRT", "ESP", "YUG")) %>%
  filter(countrycode %in% c("BRA", "MEX", "YUG", "BGR", "POL")) %>%
  filter(year >= 1950 & year <= 1990)
```

Plot the GDP per capita from 1945-1990

```{r}
shapes <- c(0,1,2,5,6)
comp_data %>%
  ggplot(aes(x = year, y = rgdpnapc, color = country, shape = country)) +
  geom_line(size=0.4) +
  geom_point() +
  scale_color_brewer(palette = "Set1")+  
  labs(title = "GDP per capita from 1945-1990",
       x = "Year",
       y = "GDP per capita") +
  theme_tufte() +
  scale_shape_manual(values = shapes) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    )
#ggsave("images/GDP_per_capita.png", width = 10, height = 6, dpi = 300)
```

Plot the growth rate of gdp

```{r}
# growth rate = (gdp(t) - gdp(t-1)) / gdp(t-1)
comp_data %>%
  group_by(country) %>%
  mutate(growth_rate = (rgdpnapc - lag(rgdpnapc)) / lag(rgdpnapc)) %>%
  mutate(moving_avg = rollmean(growth_rate, 3, fill = NA, align = "right")) %>%
  ggplot(aes(x = year, y = moving_avg, color = country, shape = country)) +
  geom_line(size=0.4)
  geom_point()
```

## Debt

```{r}
debt <- read_csv2("data/debt.csv")
```

Clean

```{r}
colnames(debt) <- c("year", "gross_debt", "foreign_lending", "net_debt", "IMF", "IBRD", "interest_payments", "amortization", "debt_servicing", "debt_servicing_percent")

debt$debt_servicing_percent <- debt$debt_servicing_percent / 10

```

Plot Gross Debt

```{r}
debt %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y=gross_debt)) +
  geom_point(aes(y=debt_servicing)) +
  geom_line(aes(y=debt_servicing)) +
  labs(x = "Year",
       y = "Million USD"
       ) 
```

## Trade

```{r}
trade <- read_csv2("data/trade.csv")
```

combine text of type column and category column

```{r}
trade$Type <- paste(trade$Type, trade$Category, sep = "_")
trade$Category <- NULL
trade$Type <- gsub(" ", "_", trade$Type)
```

transpose the data, until now it is Type,1965,1966,..., make it Year, Type_1,Type_2, Type_3,...

```{r}
trade <- trade %>%
  t(.) %>% # transpose so that the years are in the first column
  as.data.frame(.) %>% # convert to data frame
  setNames(.[1,]) %>% # set the first row as the column names
  .[-1,] %>% # remove the first row
  rownames_to_column("year") %>% # year = index col
  mutate_all(as.numeric) %>%
  select(-c(3,5,9,18)) # remove subcategories
```

Plot Total Imports and Exports and the deficit between them

```{r}
trade %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y=IMPORTS_Total, color="red")) +
  geom_line(aes(y=EXPORTS_Total, color="blue")) +
  geom_line(aes(y=IMPORTS_Total - EXPORTS_Total, color="green")) 
```

Plot the composition of imports (all rows that begin with IMPORTS_, except the IMPORTS_Total) over all years

```{r}
trade %>%
  select(1:9) %>%
  pivot_longer(cols = -year, names_to = "type", values_to = "value") %>%
  ggplot(aes(x = year, y = value, fill = type)) +
  geom_area(position = "stack") 
```

make a percentage area plot of these

```{r}
trade %>%
  select(1:9) %>%
  pivot_longer(cols = -year, names_to = "type", values_to = "value") %>%
  group_by(year) %>%
  mutate(percent = value / sum(value) * 100) %>%
  ggplot(aes(x = year, y = percent, fill = type)) +
  geom_area(position = "stack") +
  theme_linedraw() + 
  #put legend on the graph and the corresponding sections
  theme(legend.position = "bottom") 
```

How much of the rise in imports is due to fuels and chemicals?

```{r}
trade %>%
  ggplot(aes(x = year)) +
  geom_area(aes(y = IMPORTS_Mineral_fuels+ + IMPORTS_Chemicals), fill="red") 
```

## General Facts

```{r}
facts <- read_csv("data/facts.csv")
```

```{r}
facts <- facts %>% 
  select(-Unit) %>%
  t(.) %>% # transpose so that the years are in the first column
  as.data.frame(.) %>% # convert to data frame
  setNames(.[1,]) %>% # set the first row as the column names
  .[-1,] %>% # remove the first row
  rownames_to_column("year") # year = index col
```

Rename columns

```{r}
facts <- facts %>%
  rename(
    "GNP_factor_dinar" = "Gross national product at factor cost",
    "GNP_pp_dollar" = "Gross national product per head",
    "GFAF_dinar" = "Gross fixed asset formation",
    "GFAF_percent" = "Gross fixed asset formation as percentage of GDP",
    "GNP_agriculture_percent" = "Origin of GDP - Agriculture",
    "GNP_mining_percent" = "Origin of GDP - Mining and manufacturing",
    "GNP_building_percent" = "Origin of GDP - Building",
    "GNP_other_percent" = "Origin of GDP - Other",
    "G_revenue_percent" = "General government revenue as percentage of GDP"
  ) %>%
  select(-c("Collective consumption as percentage of GDP", "Federal revenue as percentage of GDP")) %>%
  #remove all spaces in data
  mutate_all(str_replace_all, " ", "") %>%
  mutate_all(as.numeric)
```

Plot the GNP per capita

```{r}
ggplot(aes(x = year, y = GNP_pp_dollar), data=facts) +
  geom_line() +
  labs(title = "GNP per capita",
       x = "Year",
       y = "GNP per capita") 
```

Plot the gov revenue as percentage of GDP

```{r}
ggplot(aes(x = year, y = G_revenue_percent), data=facts) +
  geom_line() +
  labs(title = "General government revenue as percentage of GDP",
       x = "Year",
       y = "Percentage of GDP") 
```

## Exchange Rate

```{r}
exchange <- read_csv("data/exchange_rate.csv")
```

```{r}
exchange <- exchange %>%
  t(.) %>%
  as.data.frame(.) %>%
  rownames_to_column("year") %>%
  setNames(.[1,]) %>%
  mutate_all(as.numeric) %>%
  remove_missing()
```

```{r}
exchange %>%
  filter(Year <=1990) %>%
  ggplot(aes(x = Year, y = log(Rate))) +
  geom_line()
```

Plot the yearly change in the exchange rate

```{r}
exchange %>%
  filter(Year <=1987) %>%
  mutate(change = Rate - lag(Rate)) %>%
  ggplot(aes(x = Year, y = change)) +
  geom_line()
```

## BOP

```{r}
bop <- read_csv("data/bop.csv")
```

clean

```{r}
bop <- bop %>%
  rename_all(str_replace_all, " ", "_") %>%
  rename_all(tolower) %>%
  #rename column 2 
  rename("exports" = "exports,_f.o.b." ) %>%
  rename("imports" = "imports,_c.i.f." ) %>%
  rename("remit" = "emigrant_and_worker_remittances") %>%
  mutate_all(as.numeric) %>%
  mutate(imports = abs(imports)) %>%
  # multiply all by 1000000
  mutate_all(~ . * 1000000) %>%
  mutate(year = year / 1000000)
  #remove duplicate year 1979
```

Trade Balance

```{r}
bop %>%
  ggplot(aes(x = year)) +
  geom_col(aes(y = -exports, fill = "red")) + 
  geom_col(aes(y = imports, fill = "blue")) +
  geom_line(aes(y = trade_balance), color = "black") 
```

Current Account Balance

```{r}
bop %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = trade_balance), color = "black") +
  geom_line(aes(y = current_balance), color = "green") 
  #geom_col(aes(y = remit), fill = "blue")
```

## Migration

```{r}
migration <- read_csv("data/migration.csv")
```
Plot the number of migrants and net migration

```{r}
ggplot(migration, aes(x=year))+
  geom_line(aes(y=nr_emigration/1000,color="Nr. of Emigrants")) +
  geom_point(aes(y=nr_emigration/1000), color="blue", shape=5) +
  geom_line(aes(y=net/1000, color="Net Migration"))+
  geom_point(aes(y=net/1000), color="red", shape=6) +
  #styling
  scale_color_brewer(palette = "Set1")+  
  labs(
    title = "Yugoslav Migration 1968-1985",
    x = "Year",
    y = "Migrants in Thousands",
    ) +
  geom_hline(yintercept=0, color="grey")+
  theme_tufte() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    ) +
  scale_y_continuous(n.breaks=6)+
  xlim(c(1967,1985)) 
#ggsave("images/migration.png", width=10, height=6, dpi=300)
```


# Combine some Data

```{r}
comb <- comp_data %>%
  filter(countrycode == "YUG") %>%
  select(rgdpnapc, pop, year) %>%
  mutate(gdp_dollar = rgdpnapc * pop*100) %>%
  left_join(trade, by = "year") %>%
  left_join(debt, by = "year") %>%
  left_join(facts %>% select(year, GNP_pp_dollar), by = "year") %>%
  left_join(bop %>% select(year, trade_balance, current_balance, remit), by = "year")
```

```{r}
comb <- comb %>%
  mutate_at(vars(starts_with("IMPORTS_"), starts_with("EXPORTS_")), ~.*1000000) %>%
  mutate(gnp_dollar_yug = GNP_pp_dollar * pop*1000)
```

## Energy Imports as share of GDP

```{r}
imports_gdp_plot <- comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=IMPORTS_Mineral_fuels/gdp_dollar*100), color="red")  +
    geom_point(aes(y=IMPORTS_Mineral_fuels/gdp_dollar*100), shape=5, color="red")
imports_gdp_plot
```

```{r}
imports_total_plot <- comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=IMPORTS_Mineral_fuels/IMPORTS_Total*100), color="blue") +
    geom_point(aes(y=IMPORTS_Mineral_fuels/IMPORTS_Total*100), shape=5, color="blue")
imports_total_plot
```

```{r}
#imports_gdp_plot +
imports_total_plot + 
  scale_color_brewer(palette = "Set1")+  
  labs(
    x = "Year",
    #y = "% of GDP",
    y = "% of Imports"
    ) +
  theme_tufte() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black"),
    ) +
  xlim(c(1957,1990))

#ggsave("images/imports_gdp.png", width = 6, height = 6, dpi = 300)
ggsave("images/imports_total.png", width = 6, height = 6, dpi = 300)
```


### Gross Debt as % of GDP and Servicing as % of GDP

```{r}
comb %>%
  filter(year > 1960) %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=gross_debt*1000000/gdp_dollar), color="red") +
    geom_point(aes(y=debt_servicing*1000000/gdp_dollar), color="blue")
```

### Worker Remittances as % of GDP

```{r}
comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=remit/1000000), color="orange") +  #absolute
    geom_point(aes(y=remit/1000000), color="orange", shape=5)
    #geom_line(aes(y=remit/gdp_dollar*100), color="blue") #relative to GDP
```

```{r}
comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=remit/gdp_dollar*100), color="blue") #relative to GDP
    geom_point(aes(y=remit/gdp_dollar*100), color="blue", shape=5)
```

sinking share of worker remittances, absolutely numbers down as well as relative


### Trade Balance as % of GDP

```{r}
comb %>%
  ggplot(aes(x = year)) +
    geom_line(aes(y=-trade_balance/gdp_dollar*100), color="red") 
```
