# IV Regressions

## Model

First Stage:

$$
y_{ct} = \beta D_{ct} + \sum_{j=1}^p \gamma_j y_{ct-j} + \alpha_c + \delta_t + \epsilon_{ct}
$$

Second Stage: with D as IV

$$
D_{ct} = \sum_{j=1}^p \pi_j Z_{ct-j}+ \sum_{j=1}^p \phi y_{ct-j} + \Phi_c+ \mu_t+ v_{ct}
$$

-   Z = average of democracy in Region times initial regime cell = `demreg` in dataframe
-   D = democracy variable = `dem` variable
-   y = log gdp per capita = `y` variable

Estimated with 2SLS Regression

## Imports

```{r}
library(tidyverse)
library(plm)
library(stargazer)
```

```{r}
DDCGdata <- haven::read_dta("data/DDCGdata.dta")

pdf <- pdata.frame(DDCGdata, index=c("country_name", "year"))
rm(DDCGdata)

head(pdf)
```

## Estimation
Use PLM and IVReg type formula for estimation, because IVReg does not work with panel data!

### Model 1 = 1 Lag of the Instrument

```{r}
iv_model_1 <- plm(
    y ~ dem + plm::lag(y, 1:4) | 
    plm::lag(demreg,1) + plm::lag(y, 1:4),
    data=pdf,
    effect = "twoways",
    )

summary(iv_model_1)
```

### Model 2 = 4 Lags of the Instrument

preferred specification!

```{r}
iv_model_2 <- plm(
    y ~ dem + plm::lag(y, 1:4) | 
    plm::lag(demreg,1:4) + plm::lag(y, 1:4), # 4 lags of the instrument
    data=pdf,
    effect = "twoways",
    )

summary(iv_model_2)
```

### Model 3 = Dummies for Soviet Countries

```{r}
iv_model_3 <- plm(
    y ~ dem + plm::lag(y, 1:4) + sov1 + sov2 + sov3 +sov4| #soviet (satellite) dummies
    plm::lag(demreg,1:4) + plm::lag(y, 1:4) + sov1 + sov2 + sov3 +sov4,
    data=pdf,
    effect = "twoways",
    )

summary(iv_model_3)
```

### Model 4 = Regional Trends
```{r}
iv_model_4 <- plm(
    y ~ dem + plm::lag(y, 1:4) + rtrend2 + rtrend3 + rtrend4+ rtrend5 +rtrend6 +rtrend7 | #seven regions and their trends
    plm::lag(demreg,1:4) + plm::lag(y, 1:4) + rtrend2 + rtrend3 + rtrend4+ rtrend5 +rtrend6 +rtrend7,
    data=pdf,
    effect = "twoways",
    model = "within"
    )

summary(iv_model_4)
```

## Long Run Effects

```{r}
beta_hat <- coef(iv_model_1)["dem"]
gamma_hat <- (coef(iv_model_1)[2:5])

long_run_effect_1 <- beta_hat / (1 - sum(gamma_hat))
```

```{r}
beta_hat <- coef(iv_model_2)["dem"]
gamma_hat <- (coef(iv_model_2)[2:5])

long_run_effect_2 <- beta_hat / (1 - sum(gamma_hat))
```

```{r}
beta_hat <- coef(iv_model_3)["dem"]
gamma_hat <- (coef(iv_model_3)[2:5])

long_run_effect_3 <- beta_hat / (1 - sum(gamma_hat))
```

```{r}
beta_hat <- coef(iv_model_4)["dem"]
gamma_hat <- (coef(iv_model_4)[2:5])

long_run_effect_4 <- beta_hat / (1 - sum(gamma_hat))
```

```{r}
lre <- c(long_run_effect_1, long_run_effect_2, long_run_effect_3, long_run_effect_4)
lre <- round(lre, 3)
lre
```

## Short Run Effect

Calculate the short run effects fro 1-30 years:

Formula taken from Stata Code File (`Table2_final.do`, lines 32-87 + following)
$$
effect_{t} = \hat\beta + \sum_{j=1}^p \hat\gamma_j * effect_{t-j}
$$
```{r}
sre <- c()
```


### Model 1

```{r}
dem_shortrun <- coef(iv_model_1)["dem"]
lag1 <- (coef(iv_model_1)[2])
lag2 <- (coef(iv_model_1)[3])
lag3 <- (coef(iv_model_1)[4])
lag4 <- (coef(iv_model_1)[5])
```

```{r}
effect1 <- dem_shortrun
effect2 <- (effect1 * lag1) + dem_shortrun
effect3 <- (effect2 * lag1) + (effect1 * lag2) + dem_shortrun
effect4 <- (effect3 * lag1) + (effect2 * lag2) + (effect1 * lag3) + dem_shortrun

effects <- c(effect1,effect2, effect3, effect4)
```

```{r}
for (i in 5:30) {
  eff <- (effects[i-1] * lag1) + (effects[i-2] * lag2) + (effects[i-3] * lag3) + (effects[i-4] * lag4) + dem_shortrun
  effects <- c(effects, eff)
}
```

```{r}
sre <- c(sre, effects[25])
```

### Model 2

```{r}
dem_shortrun <- coef(iv_model_2)["dem"]
lag1 <- (coef(iv_model_2)[2])
lag2 <- (coef(iv_model_2)[3])
lag3 <- (coef(iv_model_2)[4])
lag4 <- (coef(iv_model_2)[5])
```

```{r}
effect1 <- dem_shortrun
effect2 <- (effect1 * lag1) + dem_shortrun
effect3 <- (effect2 * lag1) + (effect1 * lag2) + dem_shortrun
effect4 <- (effect3 * lag1) + (effect2 * lag2) + (effect1 * lag3) + dem_shortrun

effects <- c(effect1,effect2, effect3, effect4)
```

```{r}
for (i in 5:30) {
  eff <- (effects[i-1] * lag1) + (effects[i-2] * lag2) + (effects[i-3] * lag3) + (effects[i-4] * lag4) + dem_shortrun
  effects <- c(effects, eff)
}
```

```{r}
sre <- c(sre, effects[25])
```

### Model 3

```{r}
dem_shortrun <- coef(iv_model_3)["dem"]
lag1 <- (coef(iv_model_3)[2])
lag2 <- (coef(iv_model_3)[3])
lag3 <- (coef(iv_model_3)[4])
lag4 <- (coef(iv_model_3)[5])
```

```{r}
effect1 <- dem_shortrun
effect2 <- (effect1 * lag1) + dem_shortrun
effect3 <- (effect2 * lag1) + (effect1 * lag2) + dem_shortrun
effect4 <- (effect3 * lag1) + (effect2 * lag2) + (effect1 * lag3) + dem_shortrun

effects <- c(effect1,effect2, effect3, effect4)
```

```{r}
for (i in 5:30) {
  eff <- (effects[i-1] * lag1) + (effects[i-2] * lag2) + (effects[i-3] * lag3) + (effects[i-4] * lag4) + dem_shortrun
  effects <- c(effects, eff)
}
```

```{r}
sre <- c(sre, effects[25])
```

### Model 4

```{r}
dem_shortrun <- coef(iv_model_4)["dem"]
lag1 <- (coef(iv_model_4)[2])
lag2 <- (coef(iv_model_4)[3])
lag3 <- (coef(iv_model_4)[4])
lag4 <- (coef(iv_model_4)[5])
```

```{r}
effect1 <- dem_shortrun
effect2 <- (effect1 * lag1) + dem_shortrun
effect3 <- (effect2 * lag1) + (effect1 * lag2) + dem_shortrun
effect4 <- (effect3 * lag1) + (effect2 * lag2) + (effect1 * lag3) + dem_shortrun

effects <- c(effect1,effect2, effect3, effect4)
```

```{r}
for (i in 5:30) {
  eff <- (effects[i-1] * lag1) + (effects[i-2] * lag2) + (effects[i-3] * lag3) + (effects[i-4] * lag4) + dem_shortrun
  effects <- c(effects, eff)
}
```

```{r}
sre <- c(sre, effects[25])
```

```{r}
sre <- round(sre, 3)
sre
```

## Persistence

Calculate the persistence of Democracy for the different models with this Formula:

$$
D_{pers} = \sum_{j=1}^p \hat\gamma_j
$$

```{r}
pers1 <- sum(coef(iv_model_1)[2:5])
pers2 <- sum(coef(iv_model_2)[2:5])
pers3 <- sum(coef(iv_model_3)[2:5])
pers4 <- sum(coef(iv_model_4)[2:5])

pers <- c(pers1, pers2, pers3, pers4)
pers <- round(pers, 3)
pers
```

## Table

```{r}
#| output: asis
models <- list(iv_model_1, iv_model_2, iv_model_3, iv_model_4)

stargazer(models, 
          type = "html",
          title = "Effect of Democracy on (Log) GDP per Capita",
          keep = c("dem"),
          column.labels = c("1 Lag", "4 Lags", "Soviet Dummies", "Regional Trends"),
          dep.var.labels = "Dependent Variable",
          covariate.labels = c("Democracy"),
          omit.stat = c("ser", "rsq", "adj.rsq", "f"),
          add.lines = list( # add lines for persistence, long run effect and effect after 25 years
            c("Persistence: ", pers),
            c("Long run effect: ", lre),
            c("Effect after 25 years: ", sre)
            )
          )
```
Save to Tex:

```{r}
#| output: false
#| message: false
#| warning: false

stargazer(models, 
          type = "latex",
          out = "output/table_iv.tex",
          title = "Effect of Democracy on (Log) GDP per Capita",
          keep = c("dem"),
          column.labels = c("1 Lag", "4 Lags", "Soviet Dummies", "Regional Trends"),
          dep.var.labels = "Dependent Variable",
          covariate.labels = c("Democracy"),
          omit.stat = c("ser", "rsq", "adj.rsq", "f"),
          add.lines = list( # add lines for persistence, long run effect and effect after 25 years
            c("Persistence: ", pers),
            c("Long run effect: ", lre),
            c("Effect after 25 years: ", sre)
            )
          )
```