# Democracy Does Cause Growth

by

-   Daron Acemoglu (MIT)
-   James A Robinson (Chicago)
-   Suresh Naidu (Columbia)
-   Pascual Restrepo (Boston University)

## Question and Problems

Does democracy cause growth?

Problems with the Question:

-   direct comparison between democracies / non-democracies difficult
-   correlation with other changes
-   indices of democracy lacking

## Data

Import the Data

```{r}
library(tidyverse)
library(haven)
```

```{r}
df <- read_dta("DDCGdata.dta")
```

```{r}
head(df)
```

1177 Columns and around 10000 rows

```{r}
df$
```

## Regression

Regression Formulae:

$$
y_{ct} = \beta D_t + E_{j=1}^p \gamma_j y_{ct-j} + \alpha_c + \delta_t + \epsilon_{ct}
$$

Formula: ...

-   $y_ct$ = log GDP per capita in country *c* at time *t*
    -   column: `gdppercapitaconstant2000us`
-   $D_{ct}$ = Dichotomous measure of democracy
    -   column: `yeardem`
-   *p* lags of log GDP for control
-   $\alpha$ = country fixed effects
-   $e$ = error

Estimators

-   within estimators
-   HHK estimators (not implemented in R!)
-   GMM Estimation with `pdynmc` [Intro here](https://web.archive.org/web/20220216220804/https://cran.r-project.org/web/packages/pdynmc/vignettes/pdynmc-pres-in-a-nutshell.pdf)

```{r}
library(stargazer)
library(gmm)
```

Simple Dynamic Linear Panel Model

## Results
